\section{Single-jointed arm}
\label{sec:ss_model_single-jointed_arm}

This single-jointed arm consists of a DC brushed motor attached to a pulley that
spins a straight bar in pitch.
\begin{bookfigure}
  \input{figs/single-jointed-arm-system-diagram}
  \caption{Single-jointed arm system diagram}
\end{bookfigure}

\subsection{Continuous state-space model}
\index{FRC models!single-jointed arm equations}

Using equation \eqref{eq:dot_omega_arm}, the angle and angular rate derivatives
of the arm can be written as
\begin{align}
  \dot{\theta}_{arm} &= \omega_{arm} \\
  \dot{\omega}_{arm} &= -\frac{G^2 K_t}{K_v RJ} \omega_{arm} + \frac{G K_t}{RJ} V
\end{align}

Factor out $\omega_{arm}$ and $V$ into column vectors.
\begin{align*}
  \dot{\begin{bmatrix}
    \omega_{arm}
  \end{bmatrix}} &=
  \begin{bmatrix}
    -\frac{G^2 K_t}{K_v RJ}
  \end{bmatrix}
  \begin{bmatrix}
    \omega_{arm}
  \end{bmatrix} +
  \begin{bmatrix}
    \frac{GK_t}{RJ}
  \end{bmatrix}
  \begin{bmatrix}
    V
  \end{bmatrix}
  \intertext{Augment the matrix equation with the angle state $\theta_{arm}$,
    which has the model equation $\dot{\theta}_{arm} = \omega_{arm}$. The matrix
    elements corresponding to $\omega_{arm}$ will be $1$, and the others will be
    $0$ since they don't appear, so
    $\dot{\theta}_{arm} = 0\theta_{arm} + 1\omega_{arm} + 0V$. The existing rows
    will have zeroes inserted where $\theta_{arm}$ is multiplied in.}
  \dot{\begin{bmatrix}
    \theta_{arm} \\
    \omega_{arm}
  \end{bmatrix}} &=
  \begin{bmatrix}
    0 & 1 \\
    0 & -\frac{G^2 K_t}{K_v RJ}
  \end{bmatrix}
  \begin{bmatrix}
    \theta_{arm} \\
    \omega_{arm}
  \end{bmatrix} +
  \begin{bmatrix}
    0 \\
    \frac{GK_t}{RJ}
  \end{bmatrix}
  \begin{bmatrix}
    V
  \end{bmatrix}
\end{align*}
\begin{theorem}[Single-jointed arm state-space model]
  \begin{align*}
    \dot{\mat{x}} &= \mat{A} \mat{x} + \mat{B} \mat{u} \\
    \mat{y} &= \mat{C} \mat{x} + \mat{D} \mat{u}
  \end{align*}
  \begin{equation*}
    \mat{x} =
    \begin{bmatrix}
      \theta_{arm} \\
      \omega_{arm}
    \end{bmatrix} =
    \begin{bmatrix}
      \text{angle} \\
      \text{angular velocity}
    \end{bmatrix}
    \quad
    \mat{y} = \theta_{arm} = \text{angle}
    \quad
    \mat{u} = V = \text{voltage}
  \end{equation*}
  \begin{align}
    \mat{A} &=
    \begin{bmatrix}
      0 & 1 \\
      0 & -\frac{G^2 K_t}{K_v RJ}
    \end{bmatrix} \\
    \mat{B} &=
    \begin{bmatrix}
      0 \\
      \frac{G K_t}{RJ}
    \end{bmatrix} \\
    \mat{C} &=
    \begin{bmatrix}
      1 & 0
    \end{bmatrix} \\
    \mat{D} &= 0
  \end{align}
\end{theorem}

\subsection{Model augmentation}

As per subsection \ref{subsec:input_error_estimation}, we will now augment the
\gls{model} so a $u_{error}$ state is added to the \gls{control input}.

The \gls{plant} and \gls{observer} augmentations should be performed before the
\gls{model} is \glslink{discretization}{discretized}. After the \gls{controller}
gain is computed with the unaugmented discrete \gls{model}, the controller may
be augmented. Therefore, the \gls{plant} and \gls{observer} augmentations assume
a continuous \gls{model} and the \gls{controller} augmentation assumes a
discrete \gls{controller}.
\begin{equation*}
  \mat{x}_{aug} =
  \begin{bmatrix}
    \mat{x} \\
    u_{error}
  \end{bmatrix}
  \quad
  \mat{y} = \theta_{arm}
  \quad
  \mat{u} = V
\end{equation*}
\begin{equation}
  \mat{A}_{aug} =
  \begin{bmatrix}
    \mat{A} & \mat{B} \\
    \mat{0}_{1 \times 2} & 0
  \end{bmatrix}
  \quad
  \mat{B}_{aug} =
  \begin{bmatrix}
    \mat{B} \\
    0
  \end{bmatrix}
  \quad
  \mat{C}_{aug} =
  \begin{bmatrix}
    \mat{C} & 0
  \end{bmatrix}
  \quad
  \mat{D}_{aug} = \mat{D}
\end{equation}
\begin{equation}
  \mat{K}_{aug} = \begin{bmatrix}
    \mat{K} & 1
  \end{bmatrix}
  \quad
  \mat{r}_{aug} = \begin{bmatrix}
    \mat{r} \\
    0
  \end{bmatrix}
\end{equation}

This will compensate for unmodeled dynamics such as gravity or other external
loading from lifted objects. However, if only gravity compensation is desired,
a feedforward of the form $u_{ff} = V_{gravity} \cos\theta$ is preferred where
$V_{gravity}$ is the voltage required to hold the arm level with the ground and
$\theta$ is the angle of the arm with the ground.

\subsection{Gravity feedforward}

Input voltage is proportional to torque and gravity is a constant force, but the
torque applied against the motor varies according to the arm's angle. We'll use
sum of torques to find a compensating torque.

We'll model gravity as a disturbance described by $-mg$ where $m$ is the arm's
mass. To compensate for it, we want to find a torque that is equal and opposite
to the torque applied to the arm by gravity. The bottom row of the continuous
elevator model contains the angular acceleration terms, so $Bu_{ff}$ is angular
acceleration caused by the motor; $JBu_{ff}$ is the torque.
\begin{align*}
  J Bu_{ff} &= -(\mat{r}\times\mat{F}) \\
  J Bu_{ff} &= -(rF\cos\theta)
\end{align*}

Torque is usually written as $rF\sin\theta$ where $\theta$ is the angle between
the $\mat{r}$ and $\mat{F}$ vectors, but $\theta$ in this case is being measured
from the horizontal axis rather than the vertical one, so the force vector is
$\frac{\pi}{4}$ radians out of phase. Thus, an angle of $0$ results in the
maximum torque from gravity being applied rather than the minimum.

The force of gravity $mg$ is applied at the center of the arm's mass. For a
uniform beam, this is halfway down its length, or $\frac{L}{2}$ where $L$ is the
length of the arm.
\begin{align*}
  J Bu_{ff} &= -\left(\left(\frac{L}{2}\right)(-mg)\cos\theta\right) \\
  J Bu_{ff} &= mg \frac{L}{2}\cos\theta
  \intertext{$B = \frac{GK_t}{RJ}$, so}
  J \frac{GK_t}{RJ} u_{ff} &= mg \frac{L}{2}\cos\theta \\
  u_{ff} &= \frac{RJ}{JGK_t} mg \frac{L}{2}\cos\theta \\
  u_{ff} &= \frac{L}{2} \frac{Rmg}{GK_t}\cos\theta
\end{align*}

$\frac{L}{2}$ can be adjusted according to the location of the arm's center of
mass.

\subsection{Simulation}

Python Control will be used to \glslink{discretization}{discretize} the
\gls{model} and simulate it. One of the frccontrol
examples\footnote{\url{https://github.com/calcmogul/frccontrol/blob/main/examples/single_jointed_arm.py}}
creates and tests a controller for it. Figure
\ref{fig:single_jointed_arm_response} shows the closed-loop \gls{system}
response.
\begin{svg}{build/\chapterpath/single_jointed_arm_response}
  \caption{Single-jointed arm response}
  \label{fig:single_jointed_arm_response}
\end{svg}

\subsection{Implementation}

C++ and Java implementations of this single-jointed arm controller are available
online.\footnote{\url{https://github.com/wpilibsuite/allwpilib/blob/main/wpilibcExamples/src/main/cpp/examples/StateSpaceArm/cpp/Robot.cpp}}%
\footnote{\url{https://github.com/wpilibsuite/allwpilib/blob/main/wpilibjExamples/src/main/java/edu/wpi/first/wpilibj/examples/statespacearm/Robot.java}}
