\section{Miscellaneous notation}

This book works with two-dimensional matrices in the sense that they only have
rows and columns. The dimensionality of these matrices is specified by row
first, then column. For example, a matrix with two rows and three columns would
be a two-by-three matrix. A square matrix has the same number of rows as
columns. Matrices commonly use capital letters while vectors use lowercase
letters.

The matrix $\mat{I}$ is known as the identity matrix, which is a square matrix
with ones along its diagonal and zeroes elsewhere. For example
\begin{equation*}
  \begin{bmatrix}
    1 & 0 & 0 \\
    0 & 1 & 0 \\
    0 & 0 & 1
  \end{bmatrix}
\end{equation*}

The matrix denoted by $\mat{0}_{m \times n}$ is a matrix filled with zeroes with
$m$ rows and $n$ columns.

\index{matrices!transpose}
The $\T$ in $\mat{A}\T$ denotes transpose, which flips the matrix across its
diagonal such that the rows become columns and vice versa.

\index{matrices!pseudoinverse}
The $^\dagger$ in $\mat{B}^\dagger$ denotes the Moore-Penrose pseudoinverse
given by $\mat{B}^\dagger = (\mat{B}\T\mat{B})^{-1}\mat{B}\T$. The pseudoinverse
is used when the matrix is nonsquare and thus not invertible to produce a close
approximation of an inverse in the least squares sense.

\index{matrices!trace}
$\tr(\mat{A})$ denotes the trace of the square matrix $\mat{A}$, which is
defined as the sum of the elements on the main diagonal (top-left to
bottom-right).

A matrix $\mat{M}$ is positive definite if the real number
$\mat{z}\T\mat{M}\mat{z}$ is positive for every nonzero column vector $\mat{z}$.
This is true if the eigenvalues of $\mat{M}$ are all greater than zero. Table
\ref{tab:conds_for_matrix_definiteness} describes additional definiteness
configurations.
\begin{booktable}
  \begin{tabular}{|ccc|}
    \hline
    \rowcolor{headingbg}
    \textbf{Type} & \textbf{Conditions} & \textbf{Notation} \\
    \hline
    Negative definite & All eigenvalues less than zero & $\mat{M} < \mat{0}$ \\
    Negative semidefinite & All eigenvalues less than or equal to zero &
      $\mat{M} \leq \mat{0}$ \\
    Positive semidefinite & All eigenvalues greater than or equal to zero &
      $\mat{M} \geq \mat{0}$ \\
    Positive definite & All eigenvalues greater than zero &
      $\mat{M} > \mat{0}$ \\
    \hline
  \end{tabular}
  \caption{Conditions for matrix definiteness}
  \label{tab:conds_for_matrix_definiteness}
\end{booktable}
