\section{Going digital}

The complex plane discussed so far deals with continuous \glspl{system}. In
decades past, \glspl{plant} and controllers were implemented using analog
electronics, which are continuous in nature. Nowadays, microprocessors can be
used to achieve cheaper, less complex controller designs.
\glslink{discretization}{Discretization} converts the continuous model we've
worked with so far from a set of differential equations, for example

\begin{equation}
  \dot{x} = x - 3 \label{eq:differential_equ_example}
\end{equation}

to a set of difference equations, for example

\begin{equation}
  x_{k+1} = x_k + (x_k - 3) \Delta T \label{eq:difference_equ_example}
\end{equation}

where the difference equation is run a some update rate denoted by $T$,
$\Delta T$, or sometimes sloppily as $dt$. \\

While higher order terms of a differential equation are derivatives of the state
variable (e.g., $\ddot{x}$ in relation to equation
(\ref{eq:differential_equ_example})), higher order terms of a difference
equation are delayed copies of the state variable (e.g., $x_{k-1}$ with respect
to $x_k$ in equation (\ref{eq:difference_equ_example})).

\subsection{Phase loss}

However, \gls{discretization} has drawbacks. Since a microcontroller performs
discrete steps, there is phase loss introduced in the controller. Phase loss is
the reduction of phase margin (see section \ref{sec:gain_phase_margin}) that
occurs in digital implementations of feedback controllers from sampling the
continuous system at discrete time intervals. As the sample rate of the
controller decreases, the phase margin decreases rapidly and will lead to
instability if the phase margin reaches zero. Large amounts of phase loss can
make a stable controller in the continuous domain become unstable in the
discrete domain. Here are a few ways to combat this.

\begin{itemize}
  \item Run the controller with a high sample rate.
  \item Designing the controller in the analog domain with enough phase margin
    to compensate for any phase loss that occurs as part of discretization.
  \item Convert the \gls{plant} to the digital domain and design the controller
    completely in the digital domain.
\end{itemize}

\subsection{s-plane to z-plane}

Transfer functions are converted to impulse responses using the Z-transform. The
s-plane's LHP maps to the inside of a unit circle in the z-plane. Table
\ref{tab:s-plane2z-plane} contains a few common points. \\

\begin{table}
  \caption{Mapping from s-plane to z-plane}
  \renewcommand{\arraystretch}{1.3}
  \centering
  \begin{tabular}{|cc|}
    \hline
    \rowcolor{lightblue}
    \textbf{s-plane} & \textbf{z-plane} \\
    \hline
    $(0, 0)$ & $(0, 1)$ \\
    imaginary axis & edge of unit circle \\
    $(0, -\infty)$ & $(0, 0)$ \\
    \hline
  \end{tabular}
  \label{tab:s-plane2z-plane}
\end{table}

You may notice that poles can be placed at $(0, 0)$ in the z-plane. This is
known as a deadbeat controller. An $\rm N^{th}$ order deadbeat controller decays
to the \gls{reference} in N timesteps. While this sounds great, there are other
considerations like actuation effort, \gls{robustness}, and
\gls{noise immunity}. These will be discussed in detail with LQR and LQE.

\subsection{Discretization in state-space}

This process is generally done with computers, but for completeness:

\begin{align}
  \mtx{A}_d &= e^{\mtx{A}_c T} \\
  \mtx{B}_d &= \int_0^T e^{\mtx{A}_c \tau} d\tau \mtx{B}_c \\
  \mtx{C}_d &= \mtx{C}_c \\
  \mtx{D}_d &= \mtx{D}_c
\end{align}

where a subscript of $d$ denotes discrete, a subscript of $c$ denotes
continuous, and $T$ is the sample period for the discrete system.
$e^{\mtx{A}_c T}$ and others are referred to as matrix exponentials. \\

\subsubsection{Computing the matrix exponential}

Let $\mtx{X}$ be an $n \times n$ matrix. The exponential of $\mtx{X}$ denoted by
$e^{\mtx{X}}$ is the $n \times n$ matrix given by the power series below.

\begin{equation}
  e^{\mtx{X}} = \sum_{k=0}^\infty \frac{1}{k!} \mtx{X}^k
\end{equation}

where $\mtx{X}^0$ is defined to be the identity matrix $\mtx{I}$ with the same
dimensions as $\mtx{X}$.
