\section{The Taylor series}
\index{Discretization!Taylor series}
The definition for the matrix exponential and the approximations below all use
the \textit{Taylor series expansion}. The Taylor series is a method of
approximating a function like $e^t$ via the summation of weighted polynomial
terms like $t^k$. $e^t$ has the following Taylor series around $t = 0$.

\begin{equation*}
  e^t = \sum_{n = 0}^\infty \frac{t^n}{n!}
\end{equation*}

where a finite upper bound on the number of terms produces an approximation of
$e^t$. As $n$ increases, the polynomial terms increase in power and the weights
by which they are multiplied decrease. For $e^t$ and some other functions, the
Taylor series expansion equals the original function for all values of $t$ as
the number of terms approaches infinity\footnote{Functions for which their
Taylor series expansion converges to and also equals it are called analytic
functions.}. Figure \ref{fig:taylor_series} shows the Taylor series expansion of
$e^t$ around $t = 0$ for a varying number of terms.

\begin{svg}{build/code/taylor_series}
  \caption{Taylor series expansions of $e^t$ around $t = 0$ for $n$ terms}
  \label{fig:taylor_series}
\end{svg}

We'll expand the first few terms of the Taylor series expansion in equation
(\ref{eq:mat_exp}) for $\mtx{X} = \mtx{A}T$ so we can compare it with other
methods.

\begin{equation*}
  \sum_{k=0}^3 \frac{1}{k!} (\mtx{A}T)^k = \mtx{I} + \mtx{A}T +
    \frac{1}{2}\mtx{A}^2T^2 + \frac{1}{6}\mtx{A}^3T^3
\end{equation*}

Table \ref{tab:disc_approx_matrix} compares the Taylor series expansions of the
\gls{discretization} methods for the matrix case. These use a more complex
formula which we won't present here.

\begin{booktable}
  \begin{tabular}{|cll|}
    \hline
    \rowcolor{headingbg}
    \multicolumn{1}{|c}{\textbf{Method}} &
      \multicolumn{1}{c}{\textbf{Conversion}} &
      \multicolumn{1}{c|}{\textbf{Taylor series expansion}} \\
    \hline
    Zero-order hold &
      $\mtx{\Phi} = e^{\mtx{A}T}$ &
      $\mtx{\Phi} = \mtx{I} + \mtx{A}T + \frac{1}{2}\mtx{A}^2T^2 +
        \frac{1}{6}\mtx{A}^3T^3 + \ldots$ \\
    Bilinear &
      $\mtx{\Phi} =
        \left(\mtx{I} + \frac{1}{2}\mtx{A}T\right)
        \left(\mtx{I} - \frac{1}{2}\mtx{A}T\right)^{-1}$ &
      $\mtx{\Phi} = \mtx{I} + \mtx{A}T + \frac{1}{2}\mtx{A}^2T^2 +
        \frac{1}{4}\mtx{A}^3T^3 + \ldots$ \\
    Forward Euler &
      $\mtx{\Phi} = \mtx{I} + \mtx{A}T$ &
      $\mtx{\Phi} = \mtx{I} + \mtx{A}T$ \\
    Reverse Euler &
      $\mtx{\Phi} = \left(\mtx{I} - \mtx{A}T\right)^{-1}$ &
      $\mtx{\Phi} =
        \mtx{I} + \mtx{A}T + \mtx{A}^2T^2 + \mtx{A}^3T^3 + \ldots$ \\
    \hline
  \end{tabular}
  \caption{Taylor series expansions of discretization methods (matrix case).
    The zero-order hold discretization method is exact.}
  \label{tab:disc_approx_matrix}
\end{booktable}

Each of them has different stability properties. The bilinear transform
preserves the (in)stability of the continuous time \gls{system}.
