\section{Drivetrain}

\subsection{Equations of motion}

This drivetrain consists of two DC brushed motors per side which are chained
together on their respective sides and drive wheels which are assumed to be
massless.

\begin{bookfigure}
  \begin{tikzpicture}[auto, >=latex', circuit ee IEC,
                      set resistor graphic=var resistor IEC graphic]
    % \draw [help lines] (-1,-3) grid (7,4);

    % Right wheel
    \begin{scope}[xshift=5.78cm,yshift=1.83cm]
      \draw[fill=black!50] (0.2,0.68) ellipse (0.13 and 0.67);
      \draw[fill=black!50, color=black!50] (0,0) rectangle (0.2,1.35);
      \draw[fill=white] (0,0.68) ellipse (0.13 and 0.67);
      \draw (0,1.35) -- (0.2,1.35);
      \draw (0,0) -- (0.2,0);
    \end{scope}

    % Right transmission shaft
    \begin{scope}[xshift=5.32cm,yshift=2.42cm]
      \draw[fill=black!50] (0,0) rectangle (0.46,0.1);
    \end{scope}

    % Chassis
    \begin{scope}[xshift=4.44cm,yshift=2.09cm]
      \fill[fill=white] (0,0.8) -- (0,0.2) -- (0.2,0) -- (0.2,0.2)
        -- (0.98,0.2) -- (0.78,0.8) -- cycle;
      \draw (0,0.8) -- (0.78,0.8);
      \draw (0,0.8) -- (0,0.2);
      \draw (0,0.2) -- (0.2,0);
      \draw (0,0.8) -- (0.2,0.6);
      \draw (0.78,0.8) -- (0.98,0.6);
      \draw[fill=white] (0.2,0.6) rectangle (0.98,0);
    \end{scope}

    % Left transmission shaft
    \begin{scope}[xshift=4.09cm,yshift=2.42cm]
      \draw[fill=black!50] (0,0) rectangle (0.46,0.1);
    \end{scope}

    % Left wheel
    \begin{scope}[xshift=3.75cm,yshift=1.83cm]
      \draw[fill=black!50] (0.2,0.68) ellipse (0.13 and 0.67);
      \draw[fill=black!50, color=black!50] (0,0) rectangle (0.2,1.35);
      \draw[fill=white] (0,0.68) ellipse (0.13 and 0.67);
      \draw (0,1.35) -- (0.2,1.35);
      \draw (0,0) -- (0.2,0);
    \end{scope}

    % Angular velocity arrow for left wheel
    \draw[line width=0.7pt,->] (4.24,1.97) arc (-30:30:1) node[above]
      {$\omega_l$};

    % Angular velocity arrow for right wheel
    \draw[line width=0.7pt,->] (5.44,1.97) arc (-30:30:1) node[above]
      {$\omega_r$};

    % Wheel radius arrow
    \begin{scope}[xshift=3.5cm,yshift=1.83cm]
      \draw[line width=0.7pt,<->] (0,0) -- node[left] {$r$} (0,0.67);
    \end{scope}

    % Robot radius arrow
    \begin{scope}[xshift=4.65cm,yshift=1.83cm]
      \draw[line width=0.7pt,<->] (0,0) -- node[below] {$r_b$} (0.39,0);
    \end{scope}

    % Descriptions inside graphic
    \draw (4.99,2.42) node {$J$};
  \end{tikzpicture}

  \caption{Drivetrain system diagram}
  \label{fig:drivetrain}
\end{bookfigure}

From equation (\ref{eq:tau_f}) of the flywheel \gls{model} derivations

\begin{equation}
  \tau = \frac{GK_t}{R} V - \frac{G^2K_t}{K_v R} \omega
    \label{eq:drivetrain_tau}
\end{equation}

where $\tau$ is the torque applied by one wheel of the drivetrain, $G$ is the
gear ratio of the drivetrain, $K_t$ is the torque constant of the motor, $R$ is
the resistance of the motor, and $K_v$ is the angular velocity constant. Since
$\tau = rF$ and $\omega = \frac{v}{r}$ where $v$ is the velocity of a given
drivetrain side along the ground and $r$ is the drivetrain wheel radius

\begin{align*}
  (rF) = \frac{GK_t}{R} V - \frac{G^2K_t}{K_v R} \left(\frac{v}{r}\right) \\
  rF = \frac{GK_t}{R} V - \frac{G^2K_t}{K_v Rr} v \\
  F = \frac{GK_t}{Rr} V - \frac{G^2K_t}{K_v Rr^2} v \\
  F = -\frac{G^2K_t}{K_v Rr^2} v + \frac{GK_t}{Rr} V
\end{align*}

Therefore, for each side of the robot

\begin{align*}
  F_l &= -\frac{G_l^2 K_t}{K_v R r^2} v_l + \frac{G_l K_t}{Rr} V_l \\
  F_r &= -\frac{G_r^2 K_t}{K_v R r^2} v_r + \frac{G_r K_t}{Rr} V_r
\end{align*}

where the $l$ and $r$ subscripts denote the side of the robot to which each
variable corresponds.

Let $C_1 = -\frac{G_l^2 K_t}{K_v R r^2}$, $C_2 = \frac{G_l K_t}{Rr}$,
$C_3 = -\frac{G_r^2 K_t}{K_v R r^2}$, and $C_4 = \frac{G_r K_t}{Rr}$.

\begin{align}
  F_l &= C_1 v_l + C_2 V_l \label{eq:drivetrain_Fl} \\
  F_r &= C_3 v_r + C_4 V_r \label{eq:drivetrain_Fr}
\end{align}

First, find the sum of forces.

\begin{align}
  \sum F &= ma \nonumber \\
  F_l + F_r &= m \dot{v} \nonumber \\
  F_l + F_r &= m \frac{\dot{v}_l + \dot{v}_r}{2} \nonumber \\
  \frac{2}{m} (F_l + F_r) &= \dot{v}_l + \dot{v}_r \nonumber \\
  \dot{v}_l &= \frac{2}{m} (F_l + F_r) - \dot{v}_r \label{eq:drivetrain_dotv_l}
\end{align}

Next, find the sum of torques.

\begin{align*}
  \sum \tau &= J \dot{\omega} \\
  \tau_l + \tau_r &= J \left(\frac{\dot{v}_r - \dot{v}_l}{2 r_b}\right)
\end{align*}

where $r_b$ is the radius of the drivetrain.

\begin{align*}
  (-r_b F_l) + (r_b F_r) &= J \frac{\dot{v}_r - \dot{v}_l}{2 r_b} \\
  -r_b F_l + r_b F_r &= \frac{J}{2 r_b} (\dot{v}_r - \dot{v}_l) \\
  -F_l + F_r &= \frac{J}{2 r_b^2} (\dot{v}_r - \dot{v}_l) \\
  \frac{2 r_b^2}{J} (-F_l + F_r) &= \dot{v}_r - \dot{v}_l \\
  \dot{v}_r &= \dot{v}_l + \frac{2 r_b^2}{J} (-F_l + F_r)
\end{align*}

Substitute in equation (\ref{eq:drivetrain_dotv_l}) to obtain an expression for
$\dot{v}_r$.

\begin{align}
  \dot{v}_r &= \left(\frac{2}{m} (F_l + F_r) - \dot{v}_r\right) +
    \frac{2 r_b^2}{J} (-F_l + F_r) \nonumber \\
  2\dot{v}_r &= \frac{2}{m} (F_l + F_r) + \frac{2 r_b^2}{J} (-F_l + F_r)
    \nonumber \\
  \dot{v}_r &= \frac{1}{m} (F_l + F_r) + \frac{r_b^2}{J} (-F_l + F_r)
    \label{eq:drivetrain_vr_2mid} \\
  \dot{v}_r &= \frac{1}{m} F_l + \frac{1}{m} F_r - \frac{r_b^2}{J} F_l +
    \frac{r_b^2}{J} F_r \nonumber \\
  \dot{v}_r &= \left(\frac{1}{m} - \frac{r_b^2}{J}\right) F_l +
    \left(\frac{1}{m} + \frac{r_b^2}{J}\right) F_r \label{eq:drivetrain_vr_2}
\end{align}

Substitute equation (\ref{eq:drivetrain_vr_2mid}) back into equation
(\ref{eq:drivetrain_dotv_l}) to obtain an expression for $\dot{v}_l$.

\begin{align}
  \dot{v}_l &= \frac{2}{m} (F_l + F_r) - \left(\frac{1}{m} (F_l + F_r) +
    \frac{r_b^2}{J} (-F_l + F_r)\right) \nonumber \\
  \dot{v}_l &= \frac{1}{m} (F_l + F_r) - \frac{r_b^2}{J} (-F_l + F_r)
    \nonumber \\
  \dot{v}_l &= \frac{1}{m} (F_l + F_r) + \frac{r_b^2}{J} (F_l - F_r) \nonumber
    \\
  \dot{v}_l &= \frac{1}{m} F_l + \frac{1}{m} F_r + \frac{r_b^2}{J} F_l -
    \frac{r_b^2}{J} F_r \nonumber \\
  \dot{v}_l &= \left(\frac{1}{m} + \frac{r_b^2}{J}\right) F_l +
    \left(\frac{1}{m} - \frac{r_b^2}{J}\right) F_r \label{eq:drivetrain_vl_2}
\end{align}

Now, plug the expressions for $F_l$ and $F_r$ into equation
(\ref{eq:drivetrain_vr_2}).

\begin{align}
  \dot{v}_r &= \left(\frac{1}{m} - \frac{r_b^2}{J}\right) F_l +
    \left(\frac{1}{m} + \frac{r_b^2}{J}\right) F_r \nonumber \\
  \dot{v}_r &= \left(\frac{1}{m} - \frac{r_b^2}{J}\right)
    \left(C_1 v_l + C_2 V_l\right) +
    \left(\frac{1}{m} + \frac{r_b^2}{J}\right) \left(C_3 v_r + C_4 V_r\right)
    \label{eq:drivetrain_model_right}
\end{align}

Now, plug the expressions for $F_l$ and $F_r$ into equation
(\ref{eq:drivetrain_vl_2}).

\begin{align}
  \dot{v}_l &= \left(\frac{1}{m} + \frac{r_b^2}{J}\right) F_l +
    \left(\frac{1}{m} - \frac{r_b^2}{J}\right) F_r \nonumber \\
  \dot{v}_l &= \left(\frac{1}{m} + \frac{r_b^2}{J}\right)
    \left(C_1 v_l + C_2 V_l\right) +
    \left(\frac{1}{m} - \frac{r_b^2}{J}\right) \left(C_3 v_r + C_4 V_r\right)
    \label{eq:drivetrain_model_left}
\end{align}

\subsection{Calculating constants}

\subsubsection{Moment of inertia J}

We'll use empirical measurements of linear and angular acceleration to determine
$J$. First, we'll derive the equation required to perform a linear regression
using acceleration test data.

\begin{align*}
  \tau_1 &= \mtx{r} \times \mtx{F} \\
  \tau_1 &= rma
\end{align*}

where $\tau_1$ is the torque applied by a drivetrain motor during only linear
acceleration, $r$ is the wheel radius, $m$ is the robot mass, and $a$ is the
linear acceleration.

\begin{equation*}
  \tau_2 = I \alpha
\end{equation*}

where $\tau_2$ is the torque applied by a drivetrain motor during only angular
acceleration, $I$ is the moment of inertia (same as $J$), and $\alpha$ is the
angular acceleration. If a constant voltage is applied during both the linear
and angular acceleration tests, $\tau_1 = \tau_2$. Therefore,

\begin{align}
  rma &= I \alpha \nonumber \\
  a &= \frac{I}{rm} \alpha \label{eq:drivetrain_J_regression}
\end{align}

The test procedure is as follows.

\begin{enumerate}
  \item Run the drivetrain forward at a constant voltage. Record the linear
    velocity over time.
  \item Compute the linear acceleration from the linear velocity data as the
    difference between each sample divided by the time between them.
  \item Rotate the drivetrain around its center by applying the same voltage as
    the linear acceleration test with the motors driving in opposite directions.
    Record the angular velocity over time.
  \item Compute the angular acceleration from the angular velocity data as the
    difference between each sample divided by the time between them.
  \item Perform a linear regression of linear acceleration versus angular
    acceleration. The slope of this line has the form $\frac{I}{rm}$ as per
    equation (\ref{eq:drivetrain_J_regression}).
  \item Multiply the slope by $rm$ to obtain a least squares estimate of $I$.
\end{enumerate}
