\section{Output feedback}

LQR is normally used for state feedback on

\begin{align*}
  \dot{\mtx{x}} &= \mtx{A}\mtx{x} + \mtx{B}\mtx{u} \\
  \mtx{y} &= \mtx{C}\mtx{x} + \mtx{D}\mtx{u}
\end{align*}

with the cost functional

\begin{equation*}
  J = \int_0^\infty \left(\mtx{x}^T\mtx{Q}\mtx{x} +
    \mtx{u}^T\mtx{R}\mtx{u}\right) dt
\end{equation*}

However, we may not know how to select costs for some of the states, or we don't
care what certain internal states are doing. Output feedback can accomodate
this. Not only can we make our output contain a subset of states, but we can use
any other cost metric we can think of as long as it's representable as a linear
combination of the states and inputs\footnote{We'll see this later on in section
\ref{sec:implicit_model_following} when we define the cost metric as deviation
from the behavior of another model.}.

For output feedback, we want to minimize the following cost functional.

\begin{equation*}
  J = \int_0^\infty \left(\mtx{y}^T\mtx{Q}\mtx{y} +
    \mtx{u}^T\mtx{R}\mtx{u}\right) dt
\end{equation*}

Substitute in the expression for $\mtx{y}$.

\begin{equation*}
  J = \int_0^\infty \left((\mtx{C}\mtx{x} + \mtx{D}\mtx{u})^T\mtx{Q}
    (\mtx{C}\mtx{x} + \mtx{D}\mtx{u}) + \mtx{u}^T\mtx{R}\mtx{u}\right) dt
\end{equation*}

Apply the transpose to the left-hand side of the $\mtx{Q}$ term.

\begin{equation*}
  J = \int_0^\infty \left((\mtx{x}^T\mtx{C}^T + \mtx{u}^T\mtx{D}^T)\mtx{Q}
    (\mtx{C}\mtx{x} + \mtx{D}\mtx{u}) + \mtx{u}^T\mtx{R}\mtx{u}\right) dt
\end{equation*}

Factor out $\begin{bmatrix}\mtx{x} \\ \mtx{u}\end{bmatrix}^T$ from the left side
and $\begin{bmatrix}\mtx{x} \\ \mtx{u}\end{bmatrix}$ from the right side of each
term.

\begin{align*}
  J &= \int_0^\infty \left(
    \begin{bmatrix}
      \mtx{x} \\
      \mtx{u}
    \end{bmatrix}^T
    \begin{bmatrix}
      \mtx{C}^T \\
      \mtx{D}^T
    \end{bmatrix}
    \mtx{Q}
    \begin{bmatrix}
      \mtx{C} &
      \mtx{D}
    \end{bmatrix}
    \begin{bmatrix}
      \mtx{x} \\
      \mtx{u}
    \end{bmatrix} +
    \begin{bmatrix}
      \mtx{x} \\
      \mtx{u}
    \end{bmatrix}^T
    \begin{bmatrix}
      \mtx{0} & \mtx{0} \\
      \mtx{0} & \mtx{R}
    \end{bmatrix}
    \begin{bmatrix}
      \mtx{x} \\
      \mtx{u}
    \end{bmatrix}
    \right) dt \\
  J &= \int_0^\infty \left(
    \begin{bmatrix}
      \mtx{x} \\
      \mtx{u}
    \end{bmatrix}^T
    \left(
    \begin{bmatrix}
      \mtx{C}^T \\
      \mtx{D}^T
    \end{bmatrix}
    \mtx{Q}
    \begin{bmatrix}
      \mtx{C} &
      \mtx{D}
    \end{bmatrix} +
    \begin{bmatrix}
      \mtx{0} & \mtx{0} \\
      \mtx{0} & \mtx{R}
    \end{bmatrix}
    \right)
    \begin{bmatrix}
      \mtx{x} \\
      \mtx{u}
    \end{bmatrix}
    \right) dt
\end{align*}

Multiply in $\mtx{Q}$.

\begin{equation*}
  J = \int_0^\infty \left(
    \begin{bmatrix}
      \mtx{x} \\
      \mtx{u}
    \end{bmatrix}^T
    \left(
    \begin{bmatrix}
      \mtx{C}^T\mtx{Q} \\
      \mtx{D}^T\mtx{Q}
    \end{bmatrix}
    \begin{bmatrix}
      \mtx{C} &
      \mtx{D}
    \end{bmatrix} +
    \begin{bmatrix}
      \mtx{0} & \mtx{0} \\
      \mtx{0} & \mtx{R}
    \end{bmatrix}
    \right)
    \begin{bmatrix}
      \mtx{x} \\
      \mtx{u}
    \end{bmatrix}
    \right) dt
\end{equation*}

Multiply matrices in the left term together.

\begin{equation*}
  J = \int_0^\infty \left(
    \begin{bmatrix}
      \mtx{x} \\
      \mtx{u}
    \end{bmatrix}^T
    \left(
    \begin{bmatrix}
      \mtx{C}^T\mtx{Q}\mtx{C} & \mtx{C}^T\mtx{Q}\mtx{D} \\
      \mtx{D}^T\mtx{Q}\mtx{C} & \mtx{D}^T\mtx{Q}\mtx{D}
    \end{bmatrix} +
    \begin{bmatrix}
      \mtx{0} & \mtx{0} \\
      \mtx{0} & \mtx{R}
    \end{bmatrix}
    \right)
    \begin{bmatrix}
      \mtx{x} \\
      \mtx{u}
    \end{bmatrix}
    \right) dt
\end{equation*}

Add the terms together.

\begin{equation}
  J = \int_0^\infty
  \begin{bmatrix}
    \mtx{x} \\
    \mtx{u}
  \end{bmatrix}^T
  \begin{bmatrix}
    \underbrace{\mtx{C}^T\mtx{Q}\mtx{C}}_{\mtx{Q}} &
    \underbrace{\mtx{C}^T\mtx{Q}\mtx{D}}_{\mtx{N}} \\
    \underbrace{\mtx{D}^T\mtx{Q}\mtx{C}}_{\mtx{N}^T} &
    \underbrace{\mtx{D}^T\mtx{Q}\mtx{D} + \mtx{R}}_{\mtx{R}}
  \end{bmatrix}
  \begin{bmatrix}
    \mtx{x} \\
    \mtx{u}
  \end{bmatrix}
  dt
\end{equation}

Thus, output feedback can be defined as the following optimization problem.

\begin{theorem}[Linear-quadratic regulator with output feedback]
  \begin{align*}
    \min_{\mtx{u}} &\int_0^\infty
    \begin{bmatrix}
      \mtx{x} \\
      \mtx{u}
    \end{bmatrix}^T
    \begin{bmatrix}
      \underbrace{\mtx{C}^T\mtx{Q}\mtx{C}}_{\mtx{Q}} &
      \underbrace{\mtx{C}^T\mtx{Q}\mtx{D}}_{\mtx{N}} \\
      \underbrace{\mtx{D}^T\mtx{Q}\mtx{C}}_{\mtx{N}^T} &
      \underbrace{\mtx{D}^T\mtx{Q}\mtx{D} + \mtx{R}}_{\mtx{R}}
    \end{bmatrix}
    \begin{bmatrix}
      \mtx{x} \\
      \mtx{u}
    \end{bmatrix}
    dt \\
    \text{subject to } &\dot{\mtx{x}} = \mtx{A}\mtx{x} + \mtx{B}\mtx{u}
  \end{align*}

  The optimal control policy $\mtx{u}^*$ is $\mtx{K}(\mtx{r} - \mtx{y})$ where
  $\mtx{r}$ is the desired output and $\mtx{y}$ is the measured output defined
  as $\mtx{y} = \mtx{C}\mtx{x} + \mtx{D}\mtx{u}$. $\mtx{K}$ can be computed via
  the typical LQR equations.
\end{theorem}

If the output is only the whole state vector, then $\mtx{C} = \mtx{I}$,
$\mtx{D} = \mtx{0}$, and the cost functional simplifies to that of state
feedback LQR.

\begin{equation*}
  J = \int_0^\infty
  \begin{bmatrix}
    \mtx{x} \\
    \mtx{u}
  \end{bmatrix}^T
  \begin{bmatrix}
    \mtx{Q} & \mtx{0} \\
    \mtx{0} & \mtx{R}
  \end{bmatrix}
  \begin{bmatrix}
    \mtx{x} \\
    \mtx{u}
  \end{bmatrix}
  dt
\end{equation*}
