\section{State feedback with output cost}
\index{controller design!linear-quadratic regulator!state feedback with output
  cost}

LQR is normally used for state feedback on
\begin{align*}
  \mat{x}_{k+1} &= \mat{A}\mat{x}_k + \mat{B}\mat{u}_k \\
  \mat{y}_k &= \mat{C}\mat{x}_k + \mat{D}\mat{u}_k
\end{align*}

with the cost functional
\begin{equation*}
  J = \sum_{k=0}^\infty (\mat{x}_k\T\mat{Q}\mat{x}_k +
    \mat{u}_k\T\mat{R}\mat{u}_k)
\end{equation*}

However, we may not know how to select costs for some of the states, or we don't
care what certain internal states are doing. We can address this by writing the
cost functional in terms of the output vector instead of the state vector. Not
only can we make our output contain a subset of states, but we can use any other
cost metric we can think of as long as it's representable as a linear
combination of the states and inputs.\footnote{We'll see this later on in
section \ref{sec:implicit_model_following} when we define the cost metric as
deviation from the behavior of another model.}

For state feedback with an output cost, we want to minimize the following cost
functional.
\begin{align*}
  J &= \sum_{k=0}^\infty (\mat{y}_k\T\mat{Q}\mat{y}_k +
    \mat{u}_k\T\mat{R}\mat{u}_k)
  \intertext{Substitute in the expression for $\mat{y}_k$.}
  J &= \sum_{k=0}^\infty ((\mat{C}\mat{x}_k + \mat{D}\mat{u}_k)\T\mat{Q}
    (\mat{C}\mat{x}_k + \mat{D}\mat{u}_k) + \mat{u}_k\T\mat{R}\mat{u}_k)
  \intertext{Apply the transpose to the left-hand side of the $\mat{Q}$ term.}
  J &= \sum_{k=0}^\infty ((\mat{x}_k\T\mat{C}\T + \mat{u}_k\T\mat{D}\T)\mat{Q}
    (\mat{C}\mat{x}_k + \mat{D}\mat{u}_k) + \mat{u}_k\T\mat{R}\mat{u}_k)
  \intertext{Factor out $\begin{bmatrix}\mat{x}_k \\ \mat{u}_k\end{bmatrix}\T$
    from the left side and $\begin{bmatrix}\mat{x}_k \\ \mat{u}_k\end{bmatrix}$
    from the right side of each term.}
  J &= \sum_{k=0}^\infty \left(
    \begin{bmatrix}
      \mat{x}_k \\
      \mat{u}_k
    \end{bmatrix}\T
    \begin{bmatrix}
      \mat{C}\T \\
      \mat{D}\T
    \end{bmatrix}
    \mat{Q}
    \begin{bmatrix}
      \mat{C} &
      \mat{D}
    \end{bmatrix}
    \begin{bmatrix}
      \mat{x}_k \\
      \mat{u}_k
    \end{bmatrix} +
    \begin{bmatrix}
      \mat{x}_k \\
      \mat{u}_k
    \end{bmatrix}\T
    \begin{bmatrix}
      \mat{0} & \mat{0} \\
      \mat{0} & \mat{R}
    \end{bmatrix}
    \begin{bmatrix}
      \mat{x}_k \\
      \mat{u}_k
    \end{bmatrix}
    \right) \\
  J &= \sum_{k=0}^\infty \left(
    \begin{bmatrix}
      \mat{x}_k \\
      \mat{u}_k
    \end{bmatrix}\T
    \left(
    \begin{bmatrix}
      \mat{C}\T \\
      \mat{D}\T
    \end{bmatrix}
    \mat{Q}
    \begin{bmatrix}
      \mat{C} &
      \mat{D}
    \end{bmatrix} +
    \begin{bmatrix}
      \mat{0} & \mat{0} \\
      \mat{0} & \mat{R}
    \end{bmatrix}
    \right)
    \begin{bmatrix}
      \mat{x}_k \\
      \mat{u}_k
    \end{bmatrix}
    \right)
  \intertext{Multiply in $\mat{Q}$.}
  J &= \sum_{k=0}^\infty \left(
    \begin{bmatrix}
      \mat{x}_k \\
      \mat{u}_k
    \end{bmatrix}\T
    \left(
    \begin{bmatrix}
      \mat{C}\T\mat{Q} \\
      \mat{D}\T\mat{Q}
    \end{bmatrix}
    \begin{bmatrix}
      \mat{C} &
      \mat{D}
    \end{bmatrix} +
    \begin{bmatrix}
      \mat{0} & \mat{0} \\
      \mat{0} & \mat{R}
    \end{bmatrix}
    \right)
    \begin{bmatrix}
      \mat{x}_k \\
      \mat{u}_k
    \end{bmatrix}
    \right)
  \intertext{Multiply matrices in the left term together.}
  J &= \sum_{k=0}^\infty \left(
    \begin{bmatrix}
      \mat{x}_k \\
      \mat{u}_k
    \end{bmatrix}\T
    \left(
    \begin{bmatrix}
      \mat{C}\T\mat{Q}\mat{C} & \mat{C}\T\mat{Q}\mat{D} \\
      \mat{D}\T\mat{Q}\mat{C} & \mat{D}\T\mat{Q}\mat{D}
    \end{bmatrix} +
    \begin{bmatrix}
      \mat{0} & \mat{0} \\
      \mat{0} & \mat{R}
    \end{bmatrix}
    \right)
    \begin{bmatrix}
      \mat{x}_k \\
      \mat{u}_k
    \end{bmatrix}
    \right)
\end{align*}

Add the terms together.
\begin{equation}
  J = \sum_{k=0}^\infty
  \begin{bmatrix}
    \mat{x}_k \\
    \mat{u}_k
  \end{bmatrix}\T
  \begin{bmatrix}
    \underbrace{\mat{C}\T\mat{Q}\mat{C}}_{\mat{Q}} &
    \underbrace{\mat{C}\T\mat{Q}\mat{D}}_{\mat{N}} \\
    \underbrace{\mat{D}\T\mat{Q}\mat{C}}_{\mat{N}\T} &
    \underbrace{\mat{D}\T\mat{Q}\mat{D} + \mat{R}}_{\mat{R}}
  \end{bmatrix}
  \begin{bmatrix}
    \mat{x}_k \\
    \mat{u}_k
  \end{bmatrix}
\end{equation}

Thus, state feedback with an output cost can be defined as the following
optimization problem.
\begin{theorem}[Linear-quadratic regulator with output cost]
  \begin{align}
    \min_{\mat{u}_k} &\sum_{k=0}^\infty
    \begin{bmatrix}
      \mat{x}_k \\
      \mat{u}_k
    \end{bmatrix}\T
    \begin{bmatrix}
      \underbrace{\mat{C}\T\mat{Q}\mat{C}}_{\mat{Q}} &
      \underbrace{\mat{C}\T\mat{Q}\mat{D}}_{\mat{N}} \\
      \underbrace{\mat{D}\T\mat{Q}\mat{C}}_{\mat{N}\T} &
      \underbrace{\mat{D}\T\mat{Q}\mat{D} + \mat{R}}_{\mat{R}}
    \end{bmatrix}
    \begin{bmatrix}
      \mat{x}_k \\
      \mat{u}_k
    \end{bmatrix}
    \nonumber \\
    \text{subject to } &\mat{x}_{k+1} = \mat{A}\mat{x}_k + \mat{B}\mat{u}_k
  \end{align}

  The optimal control policy $\mat{u}_k^*$ is $\mat{K}(\mat{r}_k - \mat{x}_k)$
  where $\mat{r}_k$ is the desired state. Note that the $\mat{Q}$ in
  $\mat{C}\T\mat{Q}\mat{C}$ is outputs $\times$ outputs instead of states
  $\times$ states. $\mat{K}$ can be computed via the typical LQR equations based
  on the algebraic Ricatti equation.
\end{theorem}

If the output is just the state vector, then $\mat{C} = \mat{I}$,
$\mat{D} = \mat{0}$, and the cost functional simplifies to that of LQR with a
state cost.
\begin{equation*}
  J = \sum_{k=0}^\infty
  \begin{bmatrix}
    \mat{x}_k \\
    \mat{u}_k
  \end{bmatrix}\T
  \begin{bmatrix}
    \mat{Q} & \mat{0} \\
    \mat{0} & \mat{R}
  \end{bmatrix}
  \begin{bmatrix}
    \mat{x}_k \\
    \mat{u}_k
  \end{bmatrix}
\end{equation*}
