\subsection{Transfer functions in feedback}

For \glspl{controller} to \glslink{regulator}{regulate} a \gls{system} or
\glslink{tracking}{track} a reference, they must be placed in positive or
negative feedback with the \gls{plant} (whether to use positive or negative
depends on the \gls{plant} in question). Stable feedback loops attempt to make
the \gls{output} equal the \gls{reference}.

\subsubsection{Derivation}

Given the feedback network in figure \ref{fig:closed_loop_deriv}, find an
expression for $Y(s)$.
\begin{bookfigure}
  \begin{tikzpicture}[auto, >=latex']
    % Place the blocks
    \node [name=input] {$X(s)$};
    \node [sum, right=of input] (sum) {};
    \node [block, right=of sum] (G) {$G(s)$};
    \node [right=of G] (output) {$Y(s)$};
    \node [block, below=of G] (measurements) {$H(s)$};

    % Connect the nodes
    \draw [arrow] (input) -- node[pos=0.85] {$+$} (sum);
    \draw [arrow] (sum) -- node {$Z(s)$} (G);
    \draw [arrow] (G) -- node [name=y] {} (output);
    \draw [arrow] (y) |- (measurements);
    \draw [arrow] (measurements) -| node[pos=0.99, right] {$-$} (sum);
  \end{tikzpicture}

  \caption{Closed-loop block diagram}
  \label{fig:closed_loop_deriv}
\end{bookfigure}
\begin{align}
  Y(s) &= Z(s) G(s) \nonumber \\
  Z(s) &= X(s) - Y(s) H(s) \nonumber \\
  X(s) &= Z(s) + Y(s) H(s) \nonumber \\
  X(s) &= Z(s) + Z(s) G(s) H(s) \nonumber \\
  \frac{Y(s)}{X(s)} &= \frac{Z(s) G(s)}{Z(s) + Z(s) G(s) H(s)} \nonumber \\
  \frac{Y(s)}{X(s)} &= \frac{G(s)}{1 + G(s) H(s)}
\end{align}

A more general form is
\begin{equation}
  \frac{Y(s)}{X(s)} = \frac{G(s)}{1 \mp G(s) H(s)}
    \label{eq:tf_in_feedback}
\end{equation}

where positive feedback uses the top sign and negative feedback uses the bottom
sign.

\subsubsection{Control system with feedback}
\begin{bookfigure}
  \begin{tikzpicture}[auto, >=latex']
    % Place the blocks
    \node [name=input] {$X(s)$};
    \node [sum, right=of input] (sum) {};
    \node [block, right=of sum] (K) {$K$};
    \node [block, right=of K] (G) {$G$};
    \node [right=of G] (output) {$Y(s)$};
    \node [block, below=of $(K)!0.5!(G)$] (H) {$H$};

    % Connect the nodes
    \draw [arrow] (input) -- node[pos=0.85] {$+$} (sum);
    \draw [arrow] (sum) -- node {} (K);
    \draw [arrow] (K) -- node {} (G);
    \draw [arrow] (G) -- node[name=y] {} (output);
    \draw [arrow] (y) |- (H);
    \draw [arrow] (H) -| node[pos=0.97, right] {$-$} (sum);
  \end{tikzpicture}

  \caption{Feedback controller block diagram}
  \label{fig:feedback_controller_block_diagram}
  \begin{figurekey}
    \begin{tabular}{llll}
      $X(s)$ & input & $H$ & measurement transfer function \\
      $K$ & controller gain & $Y(s)$ & output \\
      $G$ & plant transfer function & & \\
    \end{tabular}
  \end{figurekey}
\end{bookfigure}

Following equation \eqref{eq:tf_in_feedback}, the transfer function of figure
\ref{fig:feedback_controller_block_diagram}, a \gls{control system} diagram with
negative feedback, from input to output is
\begin{equation}
  G_{cl}(s) = \frac{Y(s)}{X(s)} = \frac{KG}{1 + KGH}
\end{equation}

The numerator is the \gls{open-loop gain} and the denominator is one plus the
gain around the feedback loop, which may include parts of the
\gls{open-loop gain}. As another example, the transfer function from the input
to the \gls{error} is
\begin{equation}
  G_{cl}(s) = \frac{E(s)}{X(s)} = \frac{1}{1 + KGH}
\end{equation}

The roots of the denominator of $G_{cl}(s)$ are different from those of the
open-loop transfer function $KG(s)$. These are called the closed-loop poles.

\subsubsection{DC motor transfer function}

If poles are much farther left in the LHP than the typical \gls{system} dynamics
exhibit, they can be considered negligible. Every \gls{system} has some form of
unmodeled high frequency, nonlinear dynamics, but they can be safely ignored
depending on the operating regime.

To demonstrate this, consider the transfer function for a second-order DC motor
(a CIM motor) from voltage to velocity.
\begin{equation*}
  G(s) = \frac{K}{(Js + b)(Ls + R) + K^2}
\end{equation*}

where $J = 3.2284 \times 10^{-6}$ $kg$-$m^2$, $b = 3.5077 \times 10^{-6}$
$N$-$m$-$s$, $K_e = K_t = 0.0181 \,V/rad/s$, $R = 0.0902 \,\Omega$, and
$L = 230 \times 10^{-6} \,H$.

This system is second-order because it has two poles; one corresponds to
velocity and the other corresponds to current. To make position the output
instead, we'll multiply by $\frac{1}{s}$ because position is the integral of
velocity.\footnote{This also has the effect of augmenting the system with a
position state, making it a third-order system.}
\begin{equation*}
  G(s) = \frac{K}{s((Js + b)(Ls + R) + K^2)}
\end{equation*}

Compare the step response of this \gls{system} (figure
\ref{fig:tf_highfreq_unstable_step}) with the step response of this \gls{system}
with $L$ set to zero (figure \ref{fig:tf_highfreq_stable_step}). For small
values of $K$, both \glspl{system} are stable and have nearly indistinguishable
\glspl{step response} due to the exceedingly small contribution from the fast
pole. The high frequency dynamics only cause instability for large values of $K$
that induce fast \glspl{system response}. In other words, the \glspl{system
response} of the second-order model and its first-order approximation are
similar for low frequency operating regimes.
\begin{bookfigure}
  \begin{minisvg}{2}{build/figs/highfreq_unstable_step}
    \caption{Step response of second-order DC motor plant augmented with
      position ($L = 230$ μH)}
    \label{fig:tf_highfreq_unstable_step}
  \end{minisvg}
  \hfill
  \begin{minisvg}{2}{build/figs/highfreq_stable_step}
    \caption{Step response of first-order DC motor plant augmented with position
      ($L = 0$ μH)}
    \label{fig:tf_highfreq_stable_step}
  \end{minisvg}
\end{bookfigure}

Why can't unstable poles close to the origin be ignored in the same way? The
response of high frequency stable poles decays rapidly. Unstable poles, on the
other hand, represent unstable dynamics which cause the \gls{system}
\gls{output} to grow to infinity. Regardless of how slow these unstable dynamics
are, they will eventually dominate the response.
